Algebraic filtering of surfaces from medical images with Julia

### 2019-07-08 src/background.tex

The \emph{Linear Algebraic Representation} (\textsc{lar})~\cite{Dicarlo:2014:TNL:2543138.2543294} aims to represent the \emph{chain complex}~\cite{TSAS} generated by a piecewise-linear geometric complex embedded either in 2D or in 3D. In few words, it gets a minimal characterization of geometry and topology of a cellular complex, i.e.~the embedding mapping $\mu : C_0 \to \E^d$ of 0-cells (vertices), as well a description of $(d-1)$-cells as subsets of vertices, and is able to return the whole chain complex 
\[ 
C_\bullet = (C_p, \partial_p) := 
C_3 \ 
\substack{
\delta_2 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_3 
}
\ C_2 \ 
\substack{
\delta_1 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_2 
}
\ C_1 \ 
\substack{
\delta_0 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_1 
}
\ C_0 .
\] 
and, in particular, any basis for linear chain spaces $C_p$, and any linear
boundary/coboundary map \(\partial_p\) and
\(\delta_p=\partial_{p+1}^\top\) between them. The domain of \textsc{lar} is the set of chain complexes generated by cell $d$-complexes ($2\leq d\leq 3$). The computer representations of \textsc{lar} are sparse binary matrices to represent both the operators and the chain bases. Note that in algebraic topology a $p$-chain is defined as a linear combination of $p$-cells with scalars from a field. When the scalar coefficients are from $\{-1, 0, +1\}$, a chain may represent \emph{any (oriented) subset of cells} from the cellular complex.

>>>>

The \emph{Linear Algebraic Representation} (\textsc{lar})~\cite{Dicarlo:2014:TNL:2543138.2543294} aims to represent the \emph{chain complex}~\cite{TSAS} generated by a piecewise-linear geometric complex embedded either in 2D or in 3D. In few words, it gets a minimal characterization of geometry and topology of a cellular complex, i.e.~the embedding mapping $\mu : C_0 \to \E^d$ of 0-cells (vertices), as well a description of $(d-1)$-cells as subsets of vertices, and is able to return the whole chain complex 
\[ 
C_\bullet = (C_p, \partial_p) := 
C_3 \ 
\substack{
\delta_2 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_3 
}
\ C_2 \ 
\substack{
\delta_1 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_2 
}
\ C_1 \ 
\substack{
\delta_0 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_1 
}
\ C_0 .
\] 
and, in particular, any basis for linear chain spaces $C_p$, and any linear
boundary/coboundary map \(\partial_p\) and
\(\delta_p=\partial_{p+1}^\top\) between them. The \emph{domain} of \textsc{lar} is the set of \textbf{chain complexes} generated by cell $d$-complexes ($2\leq d\leq 3$). The computer \emph{representations} of \textsc{lar} are \textbf{sparse binary matrices} to represent both the operators and the chain bases. Note that in algebraic topology a $p$-chain is defined as a linear combination of $p$-cells with scalars from a field. When the scalar coefficients are from $\{-1, 0, +1\}$, a chain may represent \emph{any (oriented) subset of cells} from the cellular complex.
